\documentclass[11pt,oneside,openany]{oblivoir}

\usepackage{kotex}
\usepackage{fontspec}
\setmainhangulfont{UnBatang}
\setsanshangulfont{UnDotum}
\setmainfont{TeX Gyre Termes}
\setsansfont{TeX Gyre Heros}
\setmonofont{DejaVu Sans Mono}

\usepackage[
  paperwidth=170mm,
  paperheight=240mm,
  inner=22mm,
  outer=18mm,
  top=22mm,
  bottom=24mm,
  headheight=14pt
]{geometry}
\AtBeginDvi{\special{papersize=170mm,240mm}}
\AddToHook{shipout/before}{\special{papersize=170mm,240mm}}
\AtBeginEnvironment{table}{\special{papersize=170mm,240mm}}
\AtBeginEnvironment{figure}{\special{papersize=170mm,240mm}}

\usepackage{amsmath,amssymb,mathtools,bm}
\usepackage{booktabs,tabularx,array,multirow,longtable}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage[most]{tcolorbox}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit,shapes.geometric}
\usepackage{hyperref}
\usepackage[nameinlink,noabbrev]{cleveref}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{caption}
\usepackage{microtype}
\usepackage{ragged2e}
\setlength{\emergencystretch}{2em}

\definecolor{rvlaBlue}{HTML}{1D4E89}
\definecolor{rvlaGray}{HTML}{F2F4F7}
\definecolor{rvlaLine}{HTML}{D0D5DD}
\definecolor{rvlaDark}{HTML}{101828}
\definecolor{rvlaWarn}{HTML}{8A4B08}

\hypersetup{
  colorlinks=true,
  linkcolor=rvlaBlue,
  citecolor=rvlaBlue,
  urlcolor=rvlaBlue,
  pdftitle={Real VLA: A Roboticist's Perspective},
  pdfauthor={Giseop Kim}
}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\sffamily\small Real VLA}
\renewcommand{\chaptermark}[1]{\markboth{\thechapter 장\ #1}{}}
\fancyhead[R]{\sffamily\small \leftmark}
\fancyfoot[C]{\sffamily\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\fancypagestyle{rvlatoc}{
  \fancyhf{}
  \fancyhead[R]{\sffamily\small CONTENTS\quad\thepage}
  \renewcommand{\headrulewidth}{0.3pt}
}

\titleformat{\chapter}[display]
  {\sffamily\bfseries\color{rvlaDark}}
  {\Large 제\thechapter 장}
  {0.5em}
  {\Huge}
\titlespacing*{\chapter}{0pt}{1.2em}{2.0em}
\titleformat{\section}
  {\sffamily\Large\bfseries\color{rvlaDark}}
  {\thesection}{0.8em}{}
\titleformat{\subsection}
  {\sffamily\large\bfseries\color{rvlaDark}}
  {\thesubsection}{0.8em}{}

\setlist[itemize]{leftmargin=2em,itemsep=0.25em,topsep=0.4em}
\setlist[enumerate]{leftmargin=2.2em,itemsep=0.25em,topsep=0.4em}
\renewcommand{\arraystretch}{1.25}
\numberwithin{section}{chapter}
\numberwithin{subsection}{section}
\numberwithin{equation}{chapter}
\makeatletter
\@addtoreset{figure}{chapter}
\@addtoreset{table}{chapter}
\makeatother
\renewcommand{\thefigure}{\thechapter.\arabic{figure}}
\renewcommand{\thetable}{\thechapter.\arabic{table}}
\AtBeginDocument{%
  \counterwithin{figure}{chapter}%
  \counterwithin{table}{chapter}%
  \renewcommand{\thefigure}{\thechapter.\arabic{figure}}%
  \renewcommand{\thetable}{\thechapter.\arabic{table}}%
}

\renewcommand{\contentsname}{Contents}
\renewcommand{\printtoctitle}[1]{%
  \vspace*{1.8em}%
  {\sffamily\bfseries\Huge\color{rvlaDark}#1\par}%
  \vspace{1.4em}%
  {\color{rvlaDark}\hrule height 0.35pt}%
  \vspace{2.8em}%
}
\setlength{\cftpartindent}{0em}
\setlength{\cftpartnumwidth}{2.0em}
\setlength{\cftchapterindent}{0em}
\setlength{\cftchapternumwidth}{2.1em}
\setlength{\cftsectionindent}{1.7em}
\setlength{\cftsectionnumwidth}{3.1em}
\setlength{\cftsubsectionindent}{3.8em}
\setlength{\cftsubsectionnumwidth}{4.2em}
\setlength{\cftbeforepartskip}{1.15em}
\setlength{\cftbeforechapterskip}{0.55em}
\setlength{\cftbeforesectionskip}{0.08em}
\renewcommand{\cftpartfont}{\normalfont\bfseries\color{rvlaBlue}}
\renewcommand{\cftpartpagefont}{\normalfont\bfseries}
\renewcommand{\cftchapterfont}{\normalfont\bfseries\color{rvlaBlue}}
\renewcommand{\cftchapterpagefont}{\normalfont\bfseries}
\renewcommand{\cftsectionfont}{\normalfont\color{rvlaBlue}}
\renewcommand{\cftsectionpagefont}{\normalfont}
\renewcommand{\cftsubsectionfont}{\normalfont\color{rvlaBlue}}
\renewcommand{\cftsubsectionpagefont}{\normalfont}
\renewcommand{\cftchapterleader}{\normalfont\cftdotfill{\cftdotsep}}
\renewcommand{\cftsectionleader}{\normalfont\cftdotfill{\cftdotsep}}
\renewcommand{\cftsubsectionleader}{\normalfont\cftdotfill{\cftdotsep}}
\newcounter{rvlatocpart}
\newcounter{rvlatocchapter}
\makeatletter
\renewcommand{\partnumberline}[1]{\stepcounter{rvlatocpart}\Roman{rvlatocpart}\hspace{1em}}
\renewcommand{\chapternumberline}[1]{\stepcounter{rvlatocchapter}\arabic{rvlatocchapter}\hspace{1em}}
\makeatother

\newtcolorbox{definitionbox}[1]{
  enhanced,
  breakable,
  colback=rvlaGray,
  colframe=rvlaBlue,
  coltitle=white,
  colbacktitle=rvlaBlue,
  title={#1},
  fonttitle=\sffamily\bfseries,
  boxrule=0.7pt,
  arc=1.2mm,
  left=1.1em,
  right=1.1em,
  top=0.8em,
  bottom=0.8em
}

\newtcolorbox{warningbox}[1]{
  enhanced,
  breakable,
  colback=orange!4,
  colframe=rvlaWarn,
  coltitle=white,
  colbacktitle=rvlaWarn,
  title={#1},
  fonttitle=\sffamily\bfseries,
  boxrule=0.7pt,
  arc=1.2mm,
  left=1.1em,
  right=1.1em,
  top=0.8em,
  bottom=0.8em
}

\newtcolorbox{casebox}[1]{
  enhanced,
  breakable,
  colback=white,
  colframe=rvlaDark,
  coltitle=white,
  colbacktitle=rvlaDark,
  title={#1},
  fonttitle=\sffamily\bfseries,
  boxrule=0.6pt,
  arc=1.2mm,
  left=1.0em,
  right=1.0em,
  top=0.75em,
  bottom=0.75em
}

\newcommand{\chapterwork}[3]{%
  \section{Further reading and exercises}
  \begin{description}[leftmargin=31mm,style=nextline,itemsep=0.45em]
    \item[Further reading] #1
    \item[Review question] #2
    \item[Design exercise] #3
  \end{description}
}

\newcommand{\obs}{\bm{o}}
\newcommand{\state}{\bm{s}}
\newcommand{\bel}{\bm{b}}
\newcommand{\act}{\bm{a}}
\newcommand{\ctrlcmd}{\bm{u}}
\newcommand{\conf}{\bm{q}}
\newcommand{\task}{\ell}
\newcommand{\policy}{\pi_\theta}
\newcommand{\controller}{\kappa}
\newcommand{\R}{\mathbb{R}}
\newcommand{\twodigits}[1]{\ifnum#1<10 0\fi\number#1}
\newcommand{\builddate}{\number\year-\twodigits{\month}-\twodigits{\day}}

\title{\sffamily\bfseries Real VLA\\{\Large 로봇공학자의 관점에서 본 Vision-Language-Action}}
\author{Giseop Kim}
\date{}

\begin{document}

\frontmatter
\maketitle
\clearpage
\thispagestyle{empty}
\vspace*{\fill}
{\sffamily\small
\noindent Author: Giseop Kim\\
\noindent Contact: gsk@dgist.ac.kr\\
\noindent Build date: \builddate\\
\noindent Format: 170mm $\times$ 240mm technical manuscript
}
\clearpage
\input{chapters/frontmatter}
\clearpage
\makeatletter
\begingroup
\hypersetup{linkcolor=rvlaBlue}
\pagestyle{rvlatoc}
\setcounter{rvlatocpart}{0}
\setcounter{rvlatocchapter}{-1}
{\let\addcontentsline\@gobblethree\tableofcontents}
\clearpage
\endgroup
\pagestyle{fancy}
\makeatother

\mainmatter
\setcounter{chapter}{-1}
\input{chapters/chapter00}
\clearpage
\part{Real VLA Stack}
\input{chapters/chapter01}
\clearpage
\input{chapters/chapter02}
\clearpage
\input{chapters/chapter03}
\clearpage
\input{chapters/chapter04}
\clearpage
\input{chapters/chapter05}
\clearpage
\part{Learning Foundations}
\input{chapters/chapter06}
\clearpage
\input{chapters/chapter07}
\clearpage
\input{chapters/chapter08}
\clearpage
\input{chapters/chapter09}
\clearpage
\part{From Planner to VLA}
\input{chapters/chapter10}
\clearpage
\input{chapters/chapter11}
\clearpage
\input{chapters/chapter12}
\clearpage
\input{chapters/chapter13}
\clearpage
\input{chapters/chapter14}
\clearpage
\input{chapters/chapter15}
\clearpage
\part{Deployment Evidence}
\input{chapters/chapter16}
\clearpage
\input{chapters/chapter17}
\clearpage
\input{chapters/chapter18}
\clearpage
\input{chapters/chapter19}
\clearpage
\part{Beyond Manipulation}
\input{chapters/chapter20}
\clearpage
\input{chapters/chapter21}
\clearpage
\input{chapters/chapter22}
\clearpage

\backmatter
\input{chapters/instructor_note}
\clearpage

\begin{thebibliography}{99}
\bibitem{kalman1960}
R. E. Kalman.
\newblock A New Approach to Linear Filtering and Prediction Problems.
\newblock Journal of Basic Engineering, 1960.

\bibitem{thrun2005}
Sebastian Thrun, Wolfram Burgard, and Dieter Fox.
\newblock Probabilistic Robotics.
\newblock MIT Press, 2005.

\bibitem{hartley2004}
Richard Hartley and Andrew Zisserman.
\newblock Multiple View Geometry in Computer Vision.
\newblock Cambridge University Press, second edition, 2004.

\bibitem{lowe2004sift}
David G. Lowe.
\newblock Distinctive Image Features from Scale-Invariant Keypoints.
\newblock International Journal of Computer Vision, 2004.

\bibitem{murartal2015orbslam}
Raul Mur-Artal, J. M. M. Montiel, and Juan D. Tardos.
\newblock ORB-SLAM: A Versatile and Accurate Monocular SLAM System.
\newblock IEEE Transactions on Robotics, 2015.

\bibitem{siciliano2009}
Bruno Siciliano, Lorenzo Sciavicco, Luigi Villani, and Giuseppe Oriolo.
\newblock Robotics: Modelling, Planning and Control.
\newblock Springer, 2009.

\bibitem{spong2006}
Mark W. Spong, Seth Hutchinson, and M. Vidyasagar.
\newblock Robot Modeling and Control.
\newblock Wiley, 2006.

\bibitem{lynch2017}
Kevin M. Lynch and Frank C. Park.
\newblock Modern Robotics: Mechanics, Planning, and Control.
\newblock Cambridge University Press, 2017.

\bibitem{khatib1987}
Oussama Khatib.
\newblock A Unified Approach for Motion and Force Control of Robot Manipulators: The Operational Space Formulation.
\newblock IEEE Journal on Robotics and Automation, 1987.

\bibitem{hogan1985}
Neville Hogan.
\newblock Impedance Control: An Approach to Manipulation.
\newblock ASME Journal of Dynamic Systems, Measurement, and Control, 1985.

\bibitem{mayne2000}
David Q. Mayne, James B. Rawlings, Christopher V. Rao, and Pierre O. M. Scokaert.
\newblock Constrained Model Predictive Control: Stability and Optimality.
\newblock Automatica, 2000.

\bibitem{ames2019}
Aaron D. Ames, Samuel Coogan, Magnus Egerstedt, Gennaro Notomista, Koushil Sreenath, and Paulo Tabuada.
\newblock Control Barrier Functions: Theory and Applications.
\newblock European Control Conference, 2019.

\bibitem{mason2001}
Matthew T. Mason.
\newblock Mechanics of Robotic Manipulation.
\newblock MIT Press, 2001.

\bibitem{lavalle2006}
Steven M. LaValle.
\newblock Planning Algorithms.
\newblock Cambridge University Press, 2006.

\bibitem{kavraki1996}
Lydia E. Kavraki, Petr Svestka, Jean-Claude Latombe, and Mark H. Overmars.
\newblock Probabilistic Roadmaps for Path Planning in High-Dimensional Configuration Spaces.
\newblock IEEE Transactions on Robotics and Automation, 1996.

\bibitem{lavalle2001rrt}
Steven M. LaValle and James J. Kuffner.
\newblock Randomized Kinodynamic Planning.
\newblock International Journal of Robotics Research, 2001.

\bibitem{kaelbling2011tamp}
Leslie Pack Kaelbling and Tomas Lozano-Perez.
\newblock Hierarchical Task and Motion Planning in the Now.
\newblock IEEE International Conference on Robotics and Automation, 2011.

\bibitem{ross2011dagger}
Stephane Ross, Geoffrey J. Gordon, and J. Andrew Bagnell.
\newblock A Reduction of Imitation Learning and Structured Prediction to No-Regret Online Learning.
\newblock AISTATS, 2011.

\bibitem{sutton2018}
Richard S. Sutton and Andrew G. Barto.
\newblock Reinforcement Learning: An Introduction.
\newblock MIT Press, second edition, 2018.

\bibitem{haarnoja2018sac}
Tuomas Haarnoja et al.
\newblock Soft Actor-Critic: Off-Policy Maximum Entropy Deep Reinforcement Learning with a Stochastic Actor.
\newblock ICML, 2018.

\bibitem{kumar2020cql}
Aviral Kumar, Aurick Zhou, George Tucker, and Sergey Levine.
\newblock Conservative Q-Learning for Offline Reinforcement Learning.
\newblock NeurIPS, 2020.

\bibitem{levine2016gps}
Sergey Levine, Chelsea Finn, Trevor Darrell, and Pieter Abbeel.
\newblock End-to-End Training of Deep Visuomotor Policies.
\newblock Journal of Machine Learning Research, 2016.

\bibitem{ebert2018visualforesight}
Frederik Ebert et al.
\newblock Visual Foresight: Model-Based Deep Reinforcement Learning for Vision-Based Robotic Control.
\newblock arXiv:1812.00568, 2018.

\bibitem{florence2021transporter}
Andy Zeng, Pete Florence, Jonathan Tompson, Stefan Welker, Jonathan Chien,
Maria Attarian, Travis Armstrong, Ivan Krasin, Dan Duong, Ayzaan Wahid,
Vikas Sindhwani, and Johnny Lee.
\newblock Transporter Networks: Rearranging the Visual World for Robotic Manipulation.
\newblock Conference on Robot Learning, 2020; PMLR, 2021. arXiv:2010.14406.

\bibitem{cliport2022}
Mohit Shridhar, Lucas Manuelli, and Dieter Fox.
\newblock CLIPort: What and Where Pathways for Robotic Manipulation.
\newblock Conference on Robot Learning, 2021; PMLR 164, 2022. arXiv:2109.12098.

\bibitem{ren2015fasterrcnn}
Shaoqing Ren, Kaiming He, Ross Girshick, and Jian Sun.
\newblock Faster R-CNN: Towards Real-Time Object Detection with Region Proposal Networks.
\newblock NeurIPS, 2015.

\bibitem{he2017maskrcnn}
Kaiming He, Georgia Gkioxari, Piotr Dollar, and Ross Girshick.
\newblock Mask R-CNN.
\newblock IEEE International Conference on Computer Vision, 2017.

\bibitem{mildenhall2020nerf}
Ben Mildenhall, Pratul P. Srinivasan, Matthew Tancik, Jonathan T. Barron, Ravi Ramamoorthi, and Ren Ng.
\newblock NeRF: Representing Scenes as Neural Radiance Fields for View Synthesis.
\newblock European Conference on Computer Vision, 2020.

\bibitem{vaswani2017}
Ashish Vaswani et al.
\newblock Attention Is All You Need.
\newblock NeurIPS, 2017.

\bibitem{dosovitskiy2021vit}
Alexey Dosovitskiy et al.
\newblock An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale.
\newblock ICLR, 2021.

\bibitem{clip2021}
Alec Radford et al.
\newblock Learning Transferable Visual Models From Natural Language Supervision.
\newblock ICML, 2021.

\bibitem{flamingo2022}
Jean-Baptiste Alayrac et al.
\newblock Flamingo: a Visual Language Model for Few-Shot Learning.
\newblock NeurIPS, 2022.

\bibitem{blip22023}
Junnan Li et al.
\newblock BLIP-2: Bootstrapping Language-Image Pre-training with Frozen Image Encoders and Large Language Models.
\newblock ICML, 2023.

\bibitem{saycan2022}
Michael Ahn et al.
\newblock Do As I Can, Not As I Say: Grounding Language in Robotic Affordances.
\newblock Conference on Robot Learning, 2022; PMLR 205, 2023. arXiv:2204.01691.

\bibitem{codeaspolicies2022}
Jacky Liang et al.
\newblock Code as Policies: Language Model Programs for Embodied Control.
\newblock arXiv:2209.07753, 2022.

\bibitem{innermonologue2022}
Wenlong Huang et al.
\newblock Inner Monologue: Embodied Reasoning through Planning with Language Models.
\newblock Conference on Robot Learning, 2022; PMLR 205, 2023. arXiv:2207.05608.

\bibitem{voxposer2023}
Wenlong Huang, Chen Wang, Ruohan Zhang, Yunzhu Li, Jiajun Wu, and Li Fei-Fei.
\newblock VoxPoser: Composable 3D Value Maps for Robotic Manipulation with Language Models.
\newblock Conference on Robot Learning, 2023; PMLR 229, 2023. arXiv:2307.05973.

\bibitem{gato2022}
Scott Reed et al.
\newblock A Generalist Agent.
\newblock arXiv:2205.06175, 2022.

\bibitem{bcz2022}
Eric Jang, Alex Irpan, Mohi Khansari, Daniel Kappler, Frederik Ebert,
Corey Lynch, Sergey Levine, and Chelsea Finn.
\newblock BC-Z: Zero-Shot Task Generalization with Robotic Imitation Learning.
\newblock Conference on Robot Learning, 2021; PMLR 164, 2022. arXiv:2202.02005.

\bibitem{rt12022}
Anthony Brohan et al.
\newblock RT-1: Robotics Transformer for Real-World Control at Scale.
\newblock Robotics: Science and Systems, 2023. arXiv:2212.06817.

\bibitem{palme2023}
Danny Driess et al.
\newblock PaLM-E: An Embodied Multimodal Language Model.
\newblock International Conference on Machine Learning, 2023; PMLR 202. arXiv:2303.03378.

\bibitem{rt22023}
Anthony Brohan et al.
\newblock RT-2: Vision-Language-Action Models Transfer Web Knowledge to Robotic Control.
\newblock arXiv:2307.15818, 2023.

\bibitem{robocat2023}
Konstantinos Bousmalis et al.
\newblock RoboCat: A Self-Improving Generalist Agent for Robotic Manipulation.
\newblock Transactions on Machine Learning Research, 2023. arXiv:2306.11706.

\bibitem{openx2023}
Open X-Embodiment Collaboration.
\newblock Open X-Embodiment: Robotic Learning Datasets and RT-X Models.
\newblock arXiv:2310.08864, 2023.

\bibitem{act2023}
Tony Z. Zhao et al.
\newblock Learning Fine-Grained Bimanual Manipulation with Low-Cost Hardware.
\newblock arXiv:2304.13705, 2023.

\bibitem{diffusionpolicy2023}
Cheng Chi et al.
\newblock Diffusion Policy: Visuomotor Policy Learning via Action Diffusion.
\newblock arXiv:2303.04137, 2023.

\bibitem{octo2024}
Octo Model Team et al.
\newblock Octo: An Open-Source Generalist Robot Policy.
\newblock arXiv:2405.12213, 2024.

\bibitem{openvla2024}
Moo Jin Kim et al.
\newblock OpenVLA: An Open-Source Vision-Language-Action Model.
\newblock Conference on Robot Learning, 2024; PMLR 270:2679--2713, 2025. arXiv:2406.09246.

\bibitem{pi02024}
Kevin Black et al.
\newblock $\pi_0$: A Vision-Language-Action Flow Model for General Robot Control.
\newblock arXiv:2410.24164, 2024.

\bibitem{fast2025}
Karl Pertsch et al.
\newblock FAST: Efficient Action Tokenization for Vision-Language-Action Models.
\newblock arXiv:2501.09747, 2025.

\bibitem{openvlaoft2025}
Moo Jin Kim, Chelsea Finn, and Percy Liang.
\newblock Fine-Tuning Vision-Language-Action Models: Optimizing Speed and Success.
\newblock arXiv:2502.19645, 2025.

\bibitem{rdt1b2024}
Songming Liu, Lingxuan Wu, Bangguo Li, Hengkai Tan, Huayu Chen,
Zhengyi Wang, Ke Xu, Hang Su, and Jun Zhu.
\newblock RDT-1B: a Diffusion Foundation Model for Bimanual Manipulation.
\newblock arXiv:2410.07864, 2024.

\bibitem{grootn12025}
NVIDIA et al.
\newblock GR00T N1: An Open Foundation Model for Generalist Humanoid Robots.
\newblock arXiv:2503.14734, 2025.

\bibitem{pi052025}
Physical Intelligence et al.
\newblock $\pi_{0.5}$: a Vision-Language-Action Model with Open-World Generalization.
\newblock arXiv:2504.16054, 2025.

\bibitem{bridgedatav22023}
Homer Walke et al.
\newblock BridgeData V2: A Dataset for Robot Learning at Scale.
\newblock arXiv:2308.12952, 2023.

\bibitem{droid2024}
Alexander Khazatsky et al.
\newblock DROID: A Large-Scale In-The-Wild Robot Manipulation Dataset.
\newblock arXiv:2403.12945, 2024.

\bibitem{libero2023}
Bo Liu et al.
\newblock LIBERO: Benchmarking Knowledge Transfer for Lifelong Robot Learning.
\newblock NeurIPS Datasets and Benchmarks, 2023. arXiv:2306.03310.

\bibitem{calvin2022}
Oier Mees, Lukas Hermann, Erick Rosete-Beas, and Wolfram Burgard.
\newblock CALVIN: A Benchmark for Language-Conditioned Policy Learning for Long-Horizon Robot Manipulation Tasks.
\newblock IEEE Robotics and Automation Letters, 2022. arXiv:2112.03227.

\bibitem{hulc2022}
Oier Mees, Lukas Hermann, and Wolfram Burgard.
\newblock What Matters in Language Conditioned Robotic Imitation Learning over Unstructured Data.
\newblock arXiv:2204.06252, 2022.

\bibitem{maniskill22023}
Jiayuan Gu et al.
\newblock ManiSkill2: A Unified Benchmark for Generalizable Manipulation Skills.
\newblock ICLR, 2023.

\bibitem{simplerenv2024}
Moo Jin Kim et al.
\newblock SimplerEnv: Simulated Manipulation Policy Evaluation Environments for Real-World Robots.
\newblock arXiv:2405.05941, 2024.

\bibitem{vlafoundry2026}
Jean Mercat, Sedrick Keh, Kushal Arora, Isabella Huang, Paarth Shah,
Haruki Nishimura, Shun Iwase, and Katherine Liu.
\newblock VLA Foundry: A Unified Framework for Training Vision-Language-Action Models.
\newblock arXiv:2604.19728, 2026.

\bibitem{robomimic2021}
Ajay Mandlekar et al.
\newblock What Matters in Learning from Offline Human Demonstrations for Robot Manipulation.
\newblock Conference on Robot Learning, 2021.

\bibitem{robosuite2020}
Yuke Zhu et al.
\newblock robosuite: A Modular Simulation Framework and Benchmark for Robot Learning.
\newblock arXiv:2009.12293, 2020.

\bibitem{shielding2018}
Mohammed Alshiekh et al.
\newblock Safe Reinforcement Learning via Shielding.
\newblock AAAI, 2018.

\bibitem{predictivesafetyfilter2021}
Kim P. Wabersich, Lukas Hewing, and Melanie N. Zeilinger.
\newblock A Predictive Safety Filter for Learning-Based Control of Constrained Nonlinear Dynamical Systems.
\newblock Automatica, 2021.

\bibitem{iso10218}
International Organization for Standardization.
\newblock ISO 10218: Robots and Robotic Devices -- Safety Requirements for Industrial Robots.
\newblock ISO, 2011.

\bibitem{isots15066}
International Organization for Standardization.
\newblock ISO/TS 15066: Robots and Robotic Devices -- Collaborative Robots.
\newblock ISO, 2016.
\end{thebibliography}

\end{document}
